Symbolic Weighted Language Models and\\ Quantitative Parsing over Infinite Alphabets

### 2021-07-18 main.tex
Révision Philippe sur les exemples

\begin{example}\label{ex:SWT}
We build a small \SWT over the alphabets $\Sigma$ and $\Deltai$ of
Examples~\ref{ex:running} and \ref{distance-time},
%weighted transducer model
with two states $q_0$ and $q_1$ that
calculates the temporal distance between an input performance in~$\Sigma^*$
and the subsequence
of the events of $\Deltai$ in a score.
%the performance $I$ and
%the score $O$ of Example~\ref{ex:running}.
%
When reading one performed event $\mu$
and the corresponding notated event $\nu$
(for instance MIDI pitch 69 and note name $\mbox{\footnotesize A4}$),
the weight value computed by the \SWT is the time distance between both,
(see Example~\ref{distance-time}), as modeled by transitions $\wei_{11}$ below.
%
If we meet the continuation symbol $'-'$
(this occurs for instance in \emph{ties}
$\musQuarter\!\!\!\mathrel{\raisebox{-1.5mm}{$\smile$}}\!\!\!\musEighth$,
or \emph{dots} \musQuarterDotted{}),
%that represents continuation
it is  skipped with no cost (transitions $\wei_{01}$ or weight $\one$).
%Likewise, the markups are ignored.
\[
\begin{array}{lclclcll}
\wei_{11}(q_0, \mu\ts{\tau}, \nu\ts{\tau'}, q_0) & = &
     \delta(\mu\ts{\tau}, \nu\ts{\tau'}) & \quad & %|\tau' - \tau| & \quad &
\wei_{11}(q_1, \mu\ts{\tau}, \nu\ts{\tau'}, q_0) & = &
     \delta(\mu\ts{\tau}, \nu\ts{\tau'}) &
     \mbox{if~} \nu \neq - \\ %|\tau' - \tau| & \quad &
\wei_{01}(q_0, \varepsilon, \mathsf{-}\ts{\tau'}, q_0) & = & \one & &
\wei_{01}(q_1, \varepsilon, \mathsf{-}\ts{\tau'}, q_0) & = & \one\\
\wei_{10}(q_0, \mu\ts{\tau}, \varepsilon, q_1) & = & \alpha & & %\multicolumn{3}{l}{\mathrm{for~all~} b \in \Delta}
\end{array}
\]
%
We also want to take performing errors into account,
since a performer could, for example, play an unwritten extra note.
%\lydia{reformulated this sentence}
%
%Reading such a note is modelled by
The transition $\wei_{10}$,
with an fixed weight value $\alpha \in \Semiring$,
switches from state $q_0$ (normal) to $q_1$ (error)
when reading an extra note $\mu$.
\philippe{Comprends pas cette phrase}
\florent{$\to$ reformulation}
The transitions in the second column below switch back to the normal state $q_0$.
% the metric computed by the \SWT is the smallest sum of point wise distances
% between dates of input and output events.
At last, we let $q_0$ be the only initial and final state, with
$\init(q_0) = \final(q_0) = \one$, and
$\init(q_1) = \final(q_1) = \zero$.
%\florent{dernière phrase utile?}
%This \SWT is capable of evaluating the differences between a score and a performance,
%all the while ensuring that performance errors are plausible.
%$\init(q_0, d, b) = \final(q_0, d, b) = \one$, and
%$\init(q_1, d, b) = \final(q_1, d, b) = \zero$,
%for all $d \in \Sigma$ and $b \in \Delta$.
\endex
\end{example}

@@

\section{SW Visibly Pushdown Automata}
\label{section:SWVPA}\label{sec:SWVPA}
The model presented in this section generalizes symbolic \VPA
(\SVPA~\cite{dAntonyAlur14SVPDA},
 generalizing themselves \VPA~\cite{AlurMadhusudan09nested} to infinite alphabets)
from Boolean semirings to arbitrary semiring weight domains.
It will compute on nested words over infinite alphabets,
associating to every such word a weight value.
Nested words are able to describe structures of labeled trees,
and in the context of parsing, they will be useful to
represent AST. % parse trees.
\florent{see \S 5 and App.A}

%\subsection{Definition}
\noindent
\label{sec:SWVPA-def}
Let $\Delta$ be a countable alphabet
%finite (large) or infinite,
that we assume partitioned into three
subsets~$\Deltai$, $\Deltac$, $\Deltar$,
whose elements are respectively called
\emph{internal}, \emph{call} and \emph{return} symbols,
following~\cite{AlurMadhusudan09nested}.
% names are  coined by application to functional program verification
% \begin{itemize}
% \item a set $\Deltai$ of \emph{internal symbols} denoted $a$,
% \item a set $\Deltac$ of \emph{call symbols} denoted $\call{a}$,
% \item a set $\Deltar$ of \emph{return symbols} denoted $\return{a}$.
% \end{itemize}
Let~$\< \Semiring, \oplus, \zero, \otimes, \one>$ be a commutative and complete semiring and let
$\bar\Phi = \< \Phii, \Phic, \Phir, \Phici, \Phicc, \Phicr>$
be a label theory over $\Semiring$
%In order to simplify notations, %and following the definition of Section~\ref{section:transducer},
%we shall write respectively
where $\Phii$, $\Phic$, $\Phir$ and~$\Phicx$ (with $\mathsf{x} \in \{ \mathsf{i}, \mathsf{c}, \mathsf{r}\}$)
stand respectively
for~$\Phi_\Deltai$, $\Phi_\Deltac$, $\Phi_\Deltar$ and~$\Phi_{\Deltac, \Delta_\mathsf{x}}$.
%
%Moreover, we extend this theory with a set $\Phii$
%of unary functions in $\Deltai \to \Semiring$,
%closed under $\oplus$ and $\otimes$.

\begin{example}\label{ex:nested-word}
Consider once more the linearized score of Example~\ref{ex:running},
with $\Deltai$ representing timed notes and continuations.
%the set of music events of the form $<\nu, \tau>$
%where $\nu$ is either a note (e.g. A4), the continuation symbol $-$, or a
%rest symbol R.
The subset $\Delta \setminus \Deltai$ is partitioned into
$\Deltac$ and $\Deltar$, containing respectively opening and closing parentheses.
%are used to encode the score structure:
%Let $\Deltac = \{\langle, \prec\}$ and  $\Deltar = \{\rangle, \succ\}$ where $\langle$
%(resp. $\rangle$)
%and $\prec$ (resp. $\succ$) correspond to start/end of measures, and start/end of
%tuplets. Then  \includegraphics[scale=0.20]{pictures/score5.png}
%is encoded with the nested word $\langle \prec 0, \frac{3}{4}, \frac{7}{8} \succ \rangle %\langle \prec 1, \frac{4}{3}, \frac{5}{3}  \succ \rangle$
The score representation $O$ in Example~\ref{ex:running}
% score: \includegraphics[scale=0.35,trim=0 5mm 0 0]{pictures/ex1.pdf},
is a nested-word of~$\Delta^*$.
%
Another example is the second candidate
\includegraphics[scale=0.35,trim=0 5mm 0 0]{pictures/ex2.pdf}
of transcription of~$I$,
linearized into
$O' =$
$\ccall{\mathsf{m}}\ts{0}$,
$\ccall{2}\ts{0}$,
$\mbox{\footnotesize A4}\ts{0}$,
$\ccall{2}\ts{\frac{1}{2}}$,
$-\ts{\frac{1}{2}}$,
$\mbox{\footnotesize B4}\ts{\frac{3}{4}}$,
$\creturn{2}\ts{1}$,
$\creturn{2}\ts{1}$,
$\ccall{\mathsf{m}}\ts{1}$,
$\ccall{3}\ts{1}$,
$\mbox{\footnotesize `C$\sharp$5`}\ts{1}$,
$\mbox{\footnotesize D5}\ts{1}$,
$\mbox{\footnotesize E5}\ts{\frac{4}{3}}$,
$\mbox{\footnotesize F5}\ts{\frac{5}{3}}$,
$\creturn{3}\ts{2}$,
$\creturn{\mathsf{m}}\ts{2}$,
$\creturn{\mathsf{m}}\ts{2}$
(see also Figure~\ref{fig:score-tree}).
The symbol between quotes $\mbox{\footnotesize `C$\sharp$5`}$
represents an \emph{appogiatura}, %(grace-note),
that is an ornemental note with theoretical duration zero.
\endex
\end{example}

\begin{definition}
A \emph{Symbolic Weighted Visibly Pushdown Automata} (\SWVPA)
over  $\Delta = \Deltai \uplus \Deltac \uplus \Deltar$, $\Semiring$ and $\bar\Phi$
is a tuple $A = \< Q, P, \init, \bar\wei, \final >$,
where $Q$ is a finite set of states,
$P$ is a finite set of stack symbols,
$\mathsf{in} : Q \to \Semiring$
(respectively $\mathsf{out} : Q \to \Semiring$)
are functions defining the weight for entering
(respectively leaving) a state,
and $\bar\wei$ is a sextuplet composed of the transition functions :
$\weii : Q \times P \times Q \to \Phici$,
$\weiei : Q \times Q \to \Phii$,
$\weic : Q \times P \times Q \times P \to \Phicc$,
$\weiec : Q \times P \times Q \to \Phic$,
$\weir : Q \times P \times Q \to \Phicr$,
$\weier : Q \times Q \to \Phir$.
%and
%$\weiex : Q \times Q \to \Phix$
%with $\mathsf{x} \in \{ \mathsf{i}, \mathsf{c}, \mathsf{r}\}$.
\end{definition}
%
Similarly as in Section~\ref{section:transducer},
we extend the above transition functions as follows
for all $q, q' \in Q$, $p \in P$,
$a \in \Deltai$,
$\call{c} \in \Deltac$,
$\return{r} \in \Deltar$,
overloading their names: % for simplicity:
\[
\begin{array}{lll}
\weii: Q \times [\Deltac \times P] \times \Deltai \times Q \to \Semiring &
\weii(q, c, p, a, q') = \eta_\mathsf{ci}(c, a) &
\mathrm{where~} \eta_\mathsf{ci} = \weii(q, p, q'),\\
%
\weiei: Q \times \Deltai \times Q \to \Semiring &
\weiei(q, a, q') = \phi_\mathsf{i}(a) &
\mathrm{where~} \phi_\mathsf{i} = \weiei(q, q').\\
%
\weic: Q \times [\Deltac \times P] \times  [\Deltac \times P] \times Q \to \Semiring &
\weic(q, c, p, \call{c'}, p', q') = \eta_\mathsf{cc}(c, \call{c'}) &
\mathrm{where~} \eta_\mathsf{cc} = \weic(q, p, p', q'),\\
%
\weiec: Q \times [\Deltac \times P] \times Q \to \Semiring &
\weiec(q, {\call{c}}, p', q') = \phi_\mathsf{c}({\call{c}}) &
\mathrm{where~} \phi_\mathsf{c} = \weiec(q, p, q').\\
%
\weir: Q \times [\Deltac \times P] \times \Deltar \times Q \to \Semiring &
\weir(q, {\call{c}},  p, {\return{r}}, q') = \eta_\mathsf{cr}({\call{c}},  {\return{r}}) &
\mathrm{where~} \eta_\mathsf{cr} = \weir(q, p, q'),\\
%
\weier: Q \times \Deltar \times Q \to \Semiring &
\weier(q, {\return{r}}, q') = \phi_\mathsf{r}({\return{r}}) &
\mathrm{where~} \phi_\mathsf{r} = \weier(q, q').\\
\end{array}
\]

@@

\begin{example} \label{ex:SWVPA}
We consider a \SWVPA over the alphabet of Example~\ref{ex:nested-word}
that expresses a weight related to the music notation,
more precisely to its structural complexity:
given a set of equivalent representations, we aim at choosing the simpler one.

Let us assume a third component in the symbols of $\Deltac$
which is the duration of the associated time interval.
It is used to check the timestamps in the symbols.
%
For instance, the following call transition starts in state
$q_{{i}/{c}}$, meaning that the current interval has number~$i$
amongst $c$ sub-intervals of duration~$\ell$ (as indicated by the stack top).
It reads a new time-division symbol~$\ccall{d}$.
The date for this symbol has to be the time $\tau$
for the previous time-division symbol read~$\ccall{c}$ (found on the stack)
plus the total duration of intervals read so far, $i \ell$, and
the new current interval duration is~$\frac{\ell}{d}$:
%
\(
\weic\bigl(q_{{i}/{c}},
           \ccall{c}\ts{\tau}\ts{\ell}, q,
           \ccall{d}\ts{\tau + i\ell}\ts{\frac{\ell}{d}}, q_{{i+1}/{c}},
           q_{{0}/{d}} \bigr) = \alpha_d
\).

>>>>

\begin{example}\label{ex:SWT}
We build a small \SWT over  alphabets $\Sigma$ and $\Deltai$
(Ex.~\ref{ex:running} and \ref{distance-time}),
%weighted transducer model
with two states $q_0$ and $q_1$, that
calculates the temporal distance between an input performance in~$\Sigma^*$
and the subsequence
of $\Deltai$ events in a score.
%the performance $I$ and
%the score $O$ of Example~\ref{ex:running}.
%
Given  a performed event $\mu$
and the corresponding notated event $\nu$
(e.g., MIDI pitch 69 and note $\mbox{\footnotesize A4}$),
the weight  computed by the \SWT is the time distance between both,
as modeled by transitions $\wei_{11}$ below.
%
The continuation symbol $'-'$
(met for instance in \emph{ties}
$\musQuarter\!\!\!\mathrel{\raisebox{-1.5mm}{$\smile$}}\!\!\!\musEighth$,
or \emph{dots} \musQuarterDotted{}),
%that represents continuation
is  skipped with no cost (transitions $\wei_{01}$).
%Likewise, the markups are ignored.
\[
\begin{array}{lclclcll}
\wei_{11}(q_0, \mu\ts{\tau}, \nu\ts{\tau'}, q_0) & = &
     \delta(\mu\ts{\tau}, \nu\ts{\tau'}) & \quad & %|\tau' - \tau| & \quad &
\wei_{11}(q_1, \mu\ts{\tau}, \nu\ts{\tau'}, q_0) & = &
     \delta(\mu\ts{\tau}, \nu\ts{\tau'}) &
     \mbox{if~} \nu \neq - \\ %|\tau' - \tau| & \quad &
\wei_{01}(q_0, \varepsilon, \mathsf{-}\ts{\tau'}, q_0) & = & \one & &
\wei_{01}(q_1, \varepsilon, \mathsf{-}\ts{\tau'}, q_0) & = & \one\\
\wei_{10}(q_0, \mu\ts{\tau}, \varepsilon, q_1) & = & \alpha & & %\multicolumn{3}{l}{\mathrm{for~all~} b \in \Delta}
\end{array}
\]
%
We also want to take performing errors into account,
since a performer could, for example, play an unwritten extra note.
%\lydia{reformulated this sentence}
%
%Reading such a note is modelled by
The transition $\wei_{10}$,
with an fixed weight value $\alpha \in \Semiring$,
switches from state $q_0$ (normal) to $q_1$ (error)
when reading an extra note $\mu$.
\philippe{Comprends pas cette phrase}
\florent{$\to$ reformulation}
The transitions in the second column below switch back to the normal state $q_0$.
% the metric computed by the \SWT is the smallest sum of point wise distances
% between dates of input and output events.
At last, we let $q_0$ be the only initial and final state, with
$\init(q_0) = \final(q_0) = \one$, and
$\init(q_1) = \final(q_1) = \zero$.
%\florent{dernière phrase utile?}
%This \SWT is capable of evaluating the differences between a score and a performance,
%all the while ensuring that performance errors are plausible.
%$\init(q_0, d, b) = \final(q_0, d, b) = \one$, and
%$\init(q_1, d, b) = \final(q_1, d, b) = \zero$,
%for all $d \in \Sigma$ and $b \in \Delta$.
\endex
\end{example}

@@

\section{SW Visibly Pushdown Automata}
\label{section:SWVPA}\label{sec:SWVPA}
The model presented now generalizes symbolic \VPA
(\SVPA~\cite{dAntonyAlur14SVPDA},
 generalizing themselves \VPA~\cite{AlurMadhusudan09nested} to infinite alphabets)
from Boolean semirings to arbitrary semiring domains.
It associates to every nested words over infinite alphabets a weight value.
Nested words can describe structures of labeled trees.
In the context of parsing, they will be useful to
represent AST. % parse trees.
\florent{see \S 5 and App.A}

%\subsection{Definition}
\noindent
\label{sec:SWVPA-def}
Let $\Delta$ be a countable alphabet
%finite (large) or infinite,
artitioned into three
subsets~$\Deltai$, $\Deltac$, $\Deltar$,
whose elements are respectively called
\emph{internal}, \emph{call} and \emph{return} symbols~\cite{AlurMadhusudan09nested}.
% names are  coined by application to functional program verification
% \begin{itemize}
% \item a set $\Deltai$ of \emph{internal symbols} denoted $a$,
% \item a set $\Deltac$ of \emph{call symbols} denoted $\call{a}$,
% \item a set $\Deltar$ of \emph{return symbols} denoted $\return{a}$.
% \end{itemize}
Let~$\< \Semiring, \oplus, \zero, \otimes, \one>$ be a commutative and complete semiring and let
$\bar\Phi = \< \Phii, \Phic, \Phir, \Phici, \Phicc, \Phicr>$
be a label theory over $\Semiring$
%In order to simplify notations, %and following the definition of Section~\ref{section:transducer},
%we shall write respectively
where $\Phii$, $\Phic$, $\Phir$ and~$\Phicx$ (with $\mathsf{x} \in \{ \mathsf{i}, \mathsf{c}, \mathsf{r}\}$)
stand respectively
for~$\Phi_\Deltai$, $\Phi_\Deltac$, $\Phi_\Deltar$ and~$\Phi_{\Deltac, \Delta_\mathsf{x}}$.
%
%Moreover, we extend this theory with a set $\Phii$
%of unary functions in $\Deltai \to \Semiring$,
%closed under $\oplus$ and $\otimes$.

\begin{example}\label{ex:nested-word}
The score representation $O$ in Example~\ref{ex:running}
is a nested-word of~$\Delta^*$.
$\Deltai$ corresponds to timed notes and continuations.
%the set of music events of the form $<\nu, \tau>$
%where $\nu$ is either a note (e.g. A4), the continuation symbol $-$, or a
%rest symbol R.
$\Deltac$ and $\Deltar$ contain respectively opening and closing parentheses.
%are used to encode the score structure:
%Let $\Deltac = \{\langle, \prec\}$ and  $\Deltar = \{\rangle, \succ\}$ where $\langle$
%(resp. $\rangle$)
%and $\prec$ (resp. $\succ$) correspond to start/end of measures, and start/end of
%tuplets. Then  \includegraphics[scale=0.20]{pictures/score5.png}
%is encoded with the nested word $\langle \prec 0, \frac{3}{4}, \frac{7}{8} \succ \rangle %\langle \prec 1, \frac{4}{3}, \frac{5}{3}  \succ \rangle$
Another example is the second candidate
\includegraphics[scale=0.35,trim=0 5mm 0 0]{pictures/ex2.pdf}
of transcription of~$I$,
linearized into
$O' =$
$\ccall{\mathsf{m}}\ts{0}$,
$\ccall{2}\ts{0}$,
$\mbox{\footnotesize A4}\ts{0}$,
$\ccall{2}\ts{\frac{1}{2}}$,
$-\ts{\frac{1}{2}}$,
$\mbox{\footnotesize B4}\ts{\frac{3}{4}}$,
$\creturn{2}\ts{1}$,
$\creturn{2}\ts{1}$,
$\ccall{\mathsf{m}}\ts{1}$,
$\ccall{3}\ts{1}$,
$\mbox{\footnotesize `C$\sharp$5`}\ts{1}$,
$\mbox{\footnotesize D5}\ts{1}$,
$\mbox{\footnotesize E5}\ts{\frac{4}{3}}$,
$\mbox{\footnotesize F5}\ts{\frac{5}{3}}$,
$\creturn{3}\ts{2}$,
$\creturn{\mathsf{m}}\ts{2}$,
$\creturn{\mathsf{m}}\ts{2}$
(see also Fig.~\ref{fig:score-tree}).
The symbol between quotes $\mbox{\footnotesize `C$\sharp$5`}$
represents an \emph{appogiatura}, %(grace-note),
i.e., an ornemental note with theoretical duration zero.
\endex
\end{example}

\begin{definition}
A \emph{Symbolic Weighted Visibly Pushdown Automata} (\SWVPA)
over  $\Delta = \Deltai \uplus \Deltac \uplus \Deltar$, $\Semiring$ and $\bar\Phi$
is a tuple $A = \< Q, P, \init, \bar\wei, \final >$,
where $Q$ is a finite set of states,
$P$ is a finite set of stack symbols,
$\mathsf{in} : Q \to \Semiring$
(respectively $\mathsf{out} : Q \to \Semiring$)
are functions defining the weight for entering
(respectively leaving) a state,
and $\bar\wei$ is a sextuplet composed of the transition functions :
$\weii : Q \times P \times Q \to \Phici$,
$\weiei : Q \times Q \to \Phii$,
$\weic : Q \times P \times Q \times P \to \Phicc$,
$\weiec : Q \times P \times Q \to \Phic$,
$\weir : Q \times P \times Q \to \Phicr$,
$\weier : Q \times Q \to \Phir$.
%and
%$\weiex : Q \times Q \to \Phix$
%with $\mathsf{x} \in \{ \mathsf{i}, \mathsf{c}, \mathsf{r}\}$.
\end{definition}
%
As in Section~\ref{section:transducer},
we extend the above transition functions as follows
for all $q, q' \in Q$, $p \in P$,
$a \in \Deltai$,
$\call{c} \in \Deltac$,
$\return{r} \in \Deltar$,
overloading their names: % for simplicity:
\[
\begin{array}{lll}
\weii: Q \times [\Deltac \times P] \times \Deltai \times Q \to \Semiring &
\weii(q, c, p, a, q') = \eta_\mathsf{ci}(c, a) &
\mathrm{where~} \eta_\mathsf{ci} = \weii(q, p, q'),\\
%
\weiei: Q \times \Deltai \times Q \to \Semiring &
\weiei(q, a, q') = \phi_\mathsf{i}(a) &
\mathrm{where~} \phi_\mathsf{i} = \weiei(q, q').\\
%
\weic: Q \times [\Deltac \times P] \times  [\Deltac \times P] \times Q \to \Semiring &
\weic(q, c, p, \call{c'}, p', q') = \eta_\mathsf{cc}(c, \call{c'}) &
\mathrm{where~} \eta_\mathsf{cc} = \weic(q, p, p', q'),\\
%
\weiec: Q \times [\Deltac \times P] \times Q \to \Semiring &
\weiec(q, {\call{c}}, p', q') = \phi_\mathsf{c}({\call{c}}) &
\mathrm{where~} \phi_\mathsf{c} = \weiec(q, p, q').\\
%
\weir: Q \times [\Deltac \times P] \times \Deltar \times Q \to \Semiring &
\weir(q, {\call{c}},  p, {\return{r}}, q') = \eta_\mathsf{cr}({\call{c}},  {\return{r}}) &
\mathrm{where~} \eta_\mathsf{cr} = \weir(q, p, q'),\\
%
\weier: Q \times \Deltar \times Q \to \Semiring &
\weier(q, {\return{r}}, q') = \phi_\mathsf{r}({\return{r}}) &
\mathrm{where~} \phi_\mathsf{r} = \weier(q, q').\\
\end{array}
\]

@@

\begin{example} \label{ex:SWVPA}
We consider a \SWVPA over the alphabet of Ex.~\ref{ex:nested-word}
that expresses a weight related to the music notation,
more precisely to its structural complexity.
Given a set of equivalent representations, we aim at choosing the simpler one.

Let us assume a third component in the symbols of $\Deltac$
\philippe{Je trouve l'exemple un peu confus. Je ne sais pas
quel est "l'associate time interval" d'un symbole. Je ne sais
pas à quoi fait référence "the following call transition". Je
n'ose pas trop toucher mais je crois qu'il faut revoir.}
which is the duration of the associated time interval.
It is used to check the timestamps in the symbols.
%
For instance, the following call transition starts in state
$q_{{i}/{c}}$, meaning that the current interval has number~$i$
amongst $c$ sub-intervals of duration~$\ell$ (as indicated by the stack top).
It reads a new time-division symbol~$\ccall{d}$.
The date for this symbol has to be the time $\tau$
for the previous time-division symbol read~$\ccall{c}$ (found on the stack)
plus the total duration of intervals read so far, $i \ell$, and
the new current interval duration is~$\frac{\ell}{d}$:
%
\(
\weic\bigl(q_{{i}/{c}},
           \ccall{c}\ts{\tau}\ts{\ell}, q,
           \ccall{d}\ts{\tau + i\ell}\ts{\frac{\ell}{d}}, q_{{i+1}/{c}},
           q_{{0}/{d}} \bigr) = \alpha_d
\).
